---
title: "p8105_hw6_ys3924"
author: "YISU"
date: "2024-11-24"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(broom)
library(purrr)
library(modelr)
```

### Problem 1

```{r, message=FALSE, warning=FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```
```{r}
set.seed(1)
create_bootstrap_sample = function(data) {
  data |> sample_frac(size = 1, replace = TRUE)
}
bootstrap_samples = tibble(bootstrap_id = 1:5000) |>
  mutate(data_sample = map(bootstrap_id, ~create_bootstrap_sample(weather_df)))
extract_bootstrap_metrics = function(data_sample) {
  model = lm(tmax ~ tmin, data = data_sample)
  r_squared = model |>
    broom::glance() |>
    dplyr::pull(r.squared)
  coefficients = coef(model)
  log_beta_product = log(coefficients[1] * coefficients[2])
  tibble(r_squared = r_squared, log_beta_product = log_beta_product)
}
bootstrap_results = bootstrap_samples |>
  mutate(
    metrics = map(data_sample, extract_bootstrap_metrics)
  ) |>
  unnest(metrics)
```

### Problem 2

```{r, warning=FALSE}
homicide = read.csv("data/homicide-data.csv") |>
  janitor::clean_names() |>
  mutate(city_state = paste(city, state, sep = ", "),
         resolved = if_else(disposition == "Closed by arrest",TRUE, FALSE)) |>
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")),
         victim_race %in% c("White", "Black")) |>
  mutate(victim_age = as.numeric(victim_age))
baltimore = homicide |>
  filter(city_state == "Baltimore, MD")
baltimore_model = glm(resolved ~ victim_age + victim_sex + victim_race,
                      data = baltimore,
                      family = binomial)
broom::tidy(baltimore_model, conf.int = TRUE) |>
  mutate(OR = exp(estimate),
         OR_low = exp(conf.low),
         OR_high = exp(conf.high)) |>
  filter(term == "victim_sexMale") |>
  select(OR, OR_low, OR_high)
```

The estimate of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed is 0.426. The corresponding CI is [0.324, 0.558].


```{r, warning=FALSE}
OR_cities = homicide |>
  group_by(city_state) |>
  nest() |>
  mutate(
    model = map(data, ~ glm(resolved ~ victim_age + victim_sex + victim_race, 
                            data = ., 
                            family = binomial())),
    results = map(model, ~ broom::tidy(., conf.int = TRUE) |>
                    mutate(
                      OR = exp(estimate),
                      OR_low = exp(conf.low),
                      OR_high = exp(conf.high)))) |>
  unnest(results) |>
  filter(term == "victim_sexMale") |>
  select(city_state, OR, OR_low, OR_high)
OR_cities |>
  arrange(OR) |>
  ggplot(aes(x = fct_reorder(city_state, OR), y = OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = OR_low, ymax = OR_high), width = 0.2) +
  coord_flip() +
  labs(
    title = "Odds Ratios for Solving Homicides by Gender",
    x = "City State",
    y = "Odds Ratio") +
  theme(axis.text.y = element_text(size = 7, hjust = 1))
```

According to above plot, cities like Albuquerque, NM, and Stockton, CA, have much higher odds ratios. On the other hand, cities like Baton Rouge, LA, and New York, NY, have lower odds ratios.

For most cities, the odds ratios look like they are below 1, which could mean that homicides with male victims are less likely to be solved than those with female victims. However, we can only be sure if the confidence intervals do not include 1.


### Problem 3

```{r}
birth_weight = read.csv("data/birthweight.csv") |>
  janitor::clean_names() |>
  mutate(
    babysex = factor(babysex, 
                     labels = c("male", "female")),
    frace = factor(frace, 
                   levels = c(1, 2, 3, 4, 8, 9),
                   labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
    mrace = factor(mrace, 
                   levels = c(1, 2, 3, 4, 8),
                   labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
    malform = factor(malform,
                     levels = c(0, 1),
                     labels = c("absent", "present"))) |>
  drop_na()
```


```{r}
model = lm(bwt ~ bhead + blength + ppbmi + ppwt + wtgain, data = birth_weight)
birth_weight |>
  add_predictions(model, var = "predicted_model") |>
  add_residuals(model, var = "residuals_model") |>
  ggplot(aes(x = predicted_model, y = residuals_model)) +
  geom_point(alpha = 0.6) +
  geom_hline(yintercept = 0, color = "red") +
  labs(title = "Residuals vs Fitted Values", x = "Fitted Values", y = "Residuals") +
  theme_minimal()
```

I would like to predict birth weight (bwt) based on maternal and pregnancy-related factors. The model uses hypothesized predictors informed by my prior knowledge in maternal health. I include bhead, blength, ppbmi, ppwt, wtgain as predictors.

The residuals are scattered around 0. There is no clear evidence of heteroscedasticity. 

```{r}
model2 = lm(bwt ~ blength + gaweeks, 
            data = birth_weight)
model3 = lm(bwt ~ bhead * blength * babysex, 
            data = birth_weight)
```


```{r}
set.seed(1)
cv_df = 
  crossv_mc(birth_weight, 100) |>
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)) |>
  mutate(
    model = map(train, \(df) lm(bwt ~ bhead + blength + ppbmi + ppwt + wtgain, data = df)),
    model2 = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
    model3 = map(train, \(df) lm(bwt ~ bhead * blength * babysex, data = df))) |> 
  mutate(
    rmse = map2_dbl(model, test, \(mod, df) rmse(model = mod, data = df)),
    rmse2 = map2_dbl(model2, test, \(mod, df) rmse(model = mod, data = df)),
    rmse3 = map2_dbl(model3, test, \(mod, df) rmse(model = mod, data = df)))
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin() +
  labs(
    title = "rmse comparison of models",
    x = "model",
    y = "rmse"
  )
```

According to above violin plot, my model (rmse) shows the lowest median rmse, which suggests better predictive accuracy, while model 2 (rmse2) has the highest median, which indicates less consistent performance. Model 3 (rmse3) falls between the previous two models. Overall, my model appears to perform the best due to its lower rmse values.
